Assimilation of conventional and satellite observations in a deep convection

### 2021-05-28 analysis/chapters/paper_ana_20181122.Rmd
Tiene forma?

---
title: Assimilation of conventional and satellite observations in a deep convection
  case during RELAMPAGO using the WRF-GSI-LETFK system.
author:
- name: Paola Belén Corrales
  email: paola.corrales@cima.fcen.uba.ar
  affiliation: Some Institute of Technology
  footnote: 1
- name: Bob Security
  email: bob@example.com
  affiliation: Another University
- name: Cat Memes
  email: cat@example.com
  affiliation: Another University
  footnote: 2
- name: Derek Zoolander
  email: derek@example.com
  affiliation: Some Institute of Technology
  footnote: 2
date: "`r Sys.Date()`"
output:
  rticles::elsevier_article: 
    keep_tex: true
abstract: |
  This is the abstract.
  It consists of two paragraphs.
journal: Atmospheric Research
layout: preprint
address:
- code: Some Institute of Technology
  address: Department, Street, City, State, Zip
- code: Another University
  address: Department, Street, City, State, Zip
bibliography: mybibfile.bib
csl: elsevier-harvard.csl
footnote:
- code: 1
  text: Corresponding Author
- code: 2
  text: Equal contribution
---

@@

experiments <- c("E4", "E5", "E6", "E7")
#Color blind friendly but E4 y E7 parecidos en Desaturated
colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E8 = "#CC3311") 
# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin/EXP/derived_data/"

@@

## Introduction

## Data and Methods

```{r dominio, fig.cap="Figure 1. a) The horizontal domain used on the simulations (black box) and the inner domain used for the experiment comparison (red box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles) . b) Locations of radiosonde launches during the experiment. The topography in meters is also shown (filled contours).", fig.height=5}

@@

dominio %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
                    breaks = seq(0, 6000, 500)) +
  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
                      name = NULL,
                      breaks = seq(0, 6000, 1000),
                      guide = guide_colorstrip(barwidth = 20,
                                               barheight = 0.8)) +
  geom_mapa() +
  geom_point(data = obs, aes(ConvertLongitude(lon), lat, 
                             color = source, shape = source), 
             size = 1, alpha = 0.8) + 
  scale_shape_manual(name = "a) Surface stations",  
                     breaks = c("Sfc - Non-official", "Sfc - Official"),
                     labels = c("ASWS","CSWS"), values = c(15, 17),
                     guide = guide_legend(override.aes = list(size = 2))) +
  scale_color_manual(name =  "a) Surface stations", 
                     values = c("Sfc - Non-official" = "#00695c", 
                                "Sfc - Official" = "#FD8002"),
                     breaks = c("Sfc - Non-official", "Sfc - Official"),
                     labels = c("ASWS","CSWS"),
                     guide = guide_legend(override.aes = list(size = 2))) +
  geom_point(data = square, aes(lon, lat), size = 0.2) +
  geom_point(data = square2, aes(lon, lat), size = 0.2, color = "#CC3311") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        legend.box = "vertical") +
  # ,
  # legend.margin = margin(),
  # legend.box.spacing = margin(0, 0, 0, 0)) +
  # 
  dominio %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
                    breaks = seq(0, 6000, 500)) +
  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
                      name = NULL,
                      breaks = seq(0, 6000, 1000),
                      guide = guide_colorstrip(barwidth = 20,
                                               barheight = 0.8)) +
  geom_mapa() +
  coord_sf(ylim = c(-35.5, -29), xlim = c(-66.5, -61.5)) +
  geom_jitter(data = lista_sondeos, aes(lon, lat, 
                                        color = periodo,
                                        shape = periodo), 
              alpha = 0.5, size = 1.5, width = 0.03, height = 0.03) +
  scale_color_brewer(palette = "Dark2",
                     guide = guide_legend(override.aes = list(size = 2,
                                                              alpha = 1))) +
  labs(color = "b) Soundings", shape = "b) Soundings") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        legend.box = "vertical") +
  # ,
  #         legend.margin = margin(),
  #         legend.box.spacing = margin(0, 0, 0, 0)) +
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
  theme(legend.position = 'bottom', legend.box = "vertical")
```

```{r ciclo, fig.cap="Diagram of analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours. The zoomed figure shows assimilation in one-hour centered window and new boundary conditions from GFSD every 6 hours."}
knitr::include_graphics(here("analysis", "figures", "analysis_cycle.pdf"))
```

```{r obs}
tribble(
  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
  "PREPBUFR", "x", "x", "x", "x",
  "AUTSFC", " ", "x", "x", "x",
  "SATWND", " ", " ", "x", "x",
  "RAD", " ", " ", " ", "x",
) %>% 
  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
      caption = "Observation types assimilated in each experiment.") %>% 
  kable_classic_2(full_width = F)
```

```{r count_obs, fig.height=6, fig.width=10}
files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E2/ANA/*/diagfiles/asim*ensmean")[1:67]

@@

files <- files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67]

@@

files <- files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E6/ANA/*/diagfiles/asim*ensmean")[1:67]

@@

p1 <- rbind(conv, aut, satwnd) %>% 
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, .(count_obs = .N), by = .(exp, lon.box, lat.box, date)] %>% 
  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lon.box, lat.box)] %>% 
  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
  geom_raster(aes(fill = obs_cycle), alpha = 0.8) +
  scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
                       guide = guide_colorbar(barwidth = 20)) +
  geom_mapa() +
  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                               E6 = "SATWND", E8 = "RAD"))) +
  labs(fill = "Mean number of\nobs per cycle") +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

p2 <- rbind(conv, aut, satwnd) %>% 
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, ":="(lev.box = cut_round(pressure, breaks = seq(50, 1050, 50)))] %>% 
  .[, .(count_obs = .N), by = .(exp, lev.box, date)] %>% 
  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lev.box)] %>% 
  ggplot(aes(lev.box, obs_cycle)) +
  geom_line(aes(color = exp)) +
  geom_point(aes(color = exp, shape = exp), size = 2) +
  scale_color_manual(values = colores_exp, 
                     labels = c(E2 = "CONV", E5 = "AUT", 
                                E6 = "SATWND", E8 = "RAD")) +
  scale_shape_manual(values = c(15, 16, 17), labels = c(E2 = "CONV", E5 = "AUT", 
                                                        E6 = "SATWND", E7 = "RAD")) +
  # geom_tile(aes(fill = obs_cycle)) +
  # scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
  #                      guide = guide_colorbar(barheight = NULL)) +
  scale_y_log10() +
  scale_x_reverse() +
  coord_flip() +
  labs(color = NULL, shape = NULL,
       x = "Pressure level (hPa)", y = "Mean number of\nobs per cycle") +
  theme_minimal() +
  theme(aspect.ratio = 2.5/1, 
        legend.position = "bottom",
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

p3 <- rbind(conv, aut, satwnd) %>% 
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, .(count_obs = .N), by = .(exp, date)] %>% 
  ggplot(aes(date, count_obs)) +
  geom_line(aes(color = exp)) +
  geom_point(aes(color = exp, shape = exp), size = 2) +
  scale_color_manual(values = colores_exp, 
                     labels = c(E2 = "CONV", E5 = "AUT", 
                                E6 = "SATWND", E7 = "RAD")) +
  scale_shape_manual(values = c(15, 16, 17), labels = c(E2 = "CONV", E5 = "AUT", 
                                                        E6 = "SATWND", E7 = "RAD")) +
  scale_date() +
  labs(y = "Number of obs per cycle", x = NULL,
       color = NULL, shape = NULL) +
  theme_minimal() +
  theme(aspect.ratio = 1/2,
        legend.position = "bottom", 
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

cowplot::plot_grid(p1, NULL, p2, p3, ncol = 3,  rel_widths = c(1.3, -0.3, 1))

>>>>

---
title: Assimilation of conventional and satellite observations in a deep convection
  case during RELAMPAGO using the WRF-GSI-LETFK system.
author:
- name: Paola Belén Corrales
  email: paola.corrales@cima.fcen.uba.ar
  affiliation: Some Institute of Technology
  footnote: 1
- name: Bob Security
  email: bob@example.com
  affiliation: Another University
- name: Cat Memes
  email: cat@example.com
  affiliation: Another University
  footnote: 2
- name: Derek Zoolander
  email: derek@example.com
  affiliation: Some Institute of Technology
  footnote: 2
date: "`r Sys.Date()`"
output:
  rticles::elsevier_article: 
    keep_tex: true
abstract: |
  This is the abstract.
  It consists of two paragraphs.
journal: Atmospheric Research
layout: 5p
address:
- code: Some Institute of Technology
  address: Department, Street, City, State, Zip
- code: Another University
  address: Department, Street, City, State, Zip
bibliography: paper_ANA.bib
csl: elsevier-harvard.csl
footnote:
- code: 1
  text: Corresponding Author
- code: 2
  text: Equal contribution
---

@@

experiments <- c("E4", "E5", "E6", "E7")
#Color blind friendly but E4 y E7 parecidos en Desaturated
colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E8 = "#CC3311") 
# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"

@@

# Introduction

Severe weather events cause significant human and material losses around the world. A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning. The north and central region of Argentina has been shown to have the highest frequency of favorable conditions for high-impact meteorological events [@brooks2003] and the highest occurrence of large hail events [@cecil2012]. This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014].

Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to several factors. On the one hand, mesoscale phenomena have limited predictability, while on the other hand, there is great difficulty in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and in the order of minutes). To solve this problem, data assimilation (DA) at the mesoscale is a topic that has gained increasing interest since it can provide appropriate initial conditions for the initialization of high-resolution forecasts [@sun2014].

DA methods can successfully be implemented by using observations with sufficient temporal and spatial resolution to characterize mesoscale atmospheric circulations. However surface observations, which directly measure the atmospheric state, have limited availability in South America. Other types of observations with relatively high spatial and temporal resolution, such as satellite radiance observations, measure only indirect quantities and are more difficult to incorporate to DA systems. 
The assimilation of conventional observations from surface stations have an important role at improving the initial conditions that lead to better forecasts. For example, @ha2014 showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S. [Agregar algún ejemplo local que no sea el RRA]

Given the sparsity of the local conventional observing network it is important to explore the potential benefits of including additional sources of observations at a regional scale. Recently Zhu et al studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels. Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the AIRS sensor in a regional application (cita: RRA, otros de Maru?). Moreover Bao et al studied the impact on temperature and humidity forecast over western USA while assimilating directly microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels due the microwave observations but an opposite impact for infrared data. Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts (Gao et al. 2016).

Having analyses with higher temporal resolution than those available from global forecasting systems (e.g. at hourly or sub-hourly) and using an increased number of regional observations that are not included in global models are some of the advantages of assimilation at these scales. Previous work has shown promising results on mesoscale DA in the region (goncalvesdegoncalves2015, Dillon 2019, Dillon 2020 submitted) using different approaches such as LETKF; (Hunt et al, 2007) and 3D-Var (). In particular, the use of ensemble-based methods have shown great potential for data assimilation in these scales, as they are able to characterize the flow-dependent evolution of the background error covariances. Moreover, the use of ensemble models also provides valuable information about the uncertainties in the forecast. Since 2016, the Argentinian National Meteorological Service (ANMS) has been running a convective-scale downscaling forecast (skabar2018). Recently a rapid-refresh ensemble-based DA system has been experimentally implemented during the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) Field Campaign [central Argentina, November-December 2018, Nesbitt 2021] with promising results [Dillon 2020 submitted]. This system assimilated data from high-resolution surface networks, AMDAR data from local aircraft flights, soundings, AIRS retrievals, high-resolution GOES-16 wind estimates, and local radar data every hour with grid spacing of 10 km. The forecast initialized from the analyses shows an overall similar performance to cold-start initializations from the Global Forecast System analysis, and even a positive impact in some cases.

One important challenge of the DA method is to assess the performance of a regional data assimilation system at representing the true atmospheric state and the skill of the forecasts initialized using the analysis as initial condition. In this sense, the observations and measurements taken during the intense observing period of RELAMPAGO are an invaluable resource to validate a regional data assimilation system.

The main objectives of this work are: to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system. The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during 22-23 November 2018 during the IOP period of the RELAMPAGO field campaign. On 22 November a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. The MCS lasts for 42 hours traveling approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 18 UTC 23 November. To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged the MSC. The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system. A multi-model ensemble is used in the data assimilation experiments allowing an evaluation of the performance of different model configurations in this particular case.

The rest of the paper is organized as follows: a description of the assimilation systems and assimilated observations is provided in Section XX, results are discussed in …. …..

# Data and Methods

The model simulations for the case of study were performed using the version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al., n.d.) model. The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa. The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25° resolution) [] every 6 hours. The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulation period.

The Local Ensemble Transform Kalman Filter (LETKF; ) implementation (V1.3) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; Hu et al. 2018) is chosen to generate the analysis. A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within +- 30 minutes of the analysis time are assimilated. Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals. For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al. 2006) was used as an observation operator to calculate model-simulated brightness temperatures.

To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localizations radius of  0.4 (in log pressure coordinates) as in [RRA] for all types of observations. A relaxation-to-prior inflation [Whitaker and Hamill] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.

We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis and random perturbations are used to generate the initial ensemble perturbations. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) with a smooth time evolution as in \citep{Necker2020,Maldonado2021}. In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. These random perturbations are also applied at the boundaries in order to keep the ensemble spread within the domain.

In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system. We use 9 different model configuration consisting on the combinations of 3 moist convection schemes  (Kain–Fritsch [], Grell–Freitas [], and Betts–Miller–Janjic []) and 3 planetary boundary layer schemes (Yonsei University Scheme [], Mellor–Yamada–Janjic Scheme [], and Mellor–Yamada Nakanishi Niino []). All ensemble members used the same land-surface model (Noah-MP, ) and microphysics (WRF single-moment 6–class squeme []), and radiation processes (RRTMG shortwave and longwave scheme []) parameterizations.

```{r dominio, fig.cap="a) The horizontal domain used on the simulations (black box) and the inner domain used for the experiment comparison (red box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles) . b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP07, orange triangles are radiosondes launched during IOP08 and purple square are radiosondes launched outside the IOPs. The topography in meters is also shown (filled contours).", out.width="110%"}

@@

dominio %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
                    breaks = seq(0, 6000, 500)) +
  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
                      name = NULL,
                      breaks = seq(0, 6000, 1000),
                      guide = guide_colorstrip(barwidth = 20,
                                               barheight = 0.5)) +
  geom_mapa() +
  geom_point(data = obs, aes(ConvertLongitude(lon), lat, 
                             color = source, shape = source), 
             size = 1, alpha = 0.8) + 
  scale_shape_manual(name = "a) Surface stations",  
                     breaks = c("Sfc - Non-official", "Sfc - Official"),
                     labels = c("ASWS","CSWS"), values = c(15, 17),
                     guide = guide_legend(override.aes = list(size = 2))) +
  scale_color_manual(name =  "a) Surface stations", 
                     values = c("Sfc - Non-official" = "#00695c", 
                                "Sfc - Official" = "#FD8002"),
                     breaks = c("Sfc - Non-official", "Sfc - Official"),
                     labels = c("ASWS","CSWS"),
                     guide = guide_legend(override.aes = list(size = 2))) +
  geom_point(data = square, aes(lon, lat), size = 0.2) +
  geom_point(data = square2, aes(lon, lat), size = 0.2, color = "#CC3311") +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom", 
        legend.box = "vertical") +
  # ,
  # legend.margin = margin(),
  # legend.box.spacing = margin(0, 0, 0, 0)) +
  # 
  dominio %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
                    breaks = seq(0, 6000, 500)) +
  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
                      name = NULL,
                      breaks = seq(0, 6000, 1000),
                      guide = guide_colorstrip(barwidth = 20,
                                               barheight = 0.5)) +
  geom_mapa() +
  coord_sf(ylim = c(-35.5, -29), xlim = c(-66.5, -61.5)) +
  geom_jitter(data = lista_sondeos, aes(lon, lat, 
                                        color = periodo,
                                        shape = periodo), 
              alpha = 0.5, size = 1.5, width = 0.03, height = 0.03) +
  scale_color_brewer(palette = "Dark2",
                     guide = guide_legend(override.aes = list(size = 2,
                                                              alpha = 1))) +
  labs(color = "b) Soundings", shape = "b) Soundings") +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom", 
        legend.box = "vertical") +
  # ,
  #         legend.margin = margin(),
  #         legend.box.spacing = margin(0, 0, 0, 0)) +
  plot_layout(ncol = 2, widths = c(4, 1), guides = "collect") +
  plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
  theme(legend.position = 'bottom', legend.box = "vertical")
```

### Observations

#### Conventional and Satellite-derived wind

Conventional and satellite-derived wind observations are part of the Global Data Assimilation System (GDAS) [] except for Automatic Surface Weather Stations (ASWS). These observations come from public and private automatic station networks over Southern South America and are part of RELAMPAGO Data Set repository (García, F., et al. 2019).

We used conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP (Keyser 2013) that consist of surface observations (land and sea) and upper-air observations from radiosondes and aircrafts (Figure \@ref(fig:dominio)). The frequency of these observations varied between 1 hour for surface stations and 24 hours for most radiosondes in the region.

Satellite-derived wind observations also come from GDAS and includes information from GOES-16 (estimated from the visible, infrared and water vapour channels) and METEOSAT 8 and 11 (visible and water vapour channels). Due to the domain covered by each of these satellites, GOES-16 is the primary source of this type of observation (99 % of this type of observation).

Table 2 lists all the observation types (i.e., pressure, temperature, specific humidity and wind) available for each source, together with their associated errors. The observation errors were specified following the GSI default configuration. In some cases, the error varies with height and it depends on the specific platform (aircraft and satellite-derived wind). In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and model-simulated observation based on the first-guess)  with a predefined threshold (also included in Table 2).

#### Satellite radiances

Satellite radiances available through the GDAS datastream which consist of infrared and microwave observations are used in this paper. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 3). Since the regional domain is located in mid-latitudes, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, radiance observations were not available in all analysis cycles and the number of observations varied significantly among cycles. This is evident in Figure 2 b) that shows the number of radiance observations assimilated at each analysis cycle as a function of pressure. In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in the period. The vertical location of each radiance was estimated as the model level at which its weighting function was maximized as calculated by the CRTM. The multispectral sensors have a good vertical distribution and are able to sense throughout the entire atmosphere. The channels whose maximum weighting function is in the stratosphere were rejected due to the low model top model (50 hPa) used in the experiments.

The channels accepted for assimilation and their associated error were defined using the default GSI configuration. The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically. Firstly a spatial data thinning is applied using a 60 km grid following [Singh et al 2016, Jones et al 2013, Lin et al 2017], where the observations to be assimilated are chosen based on their distance to the model grid points, to the observation quality (based on available data quality predictors) and on the number of available channels (from the same pixel and sensor) that passed the quality control. Also, observations over the sea or sea-ice surface are preferred from those over land or snow [cita a la guía avanzada].

The thinned observations were then bias corrected. The bias correction (BC) has an air-mass dependent and an angle-dependent component (Zhu 2014) and it is calculated as a multi-linear function of N predictors $p_i(x)$, with associated coefficients $\beta_i$. Then, the bias corrected brightness temperature can be obtained as:

$$\mathit{BT_bc}  =\mathit{ BT} +  \sum{i = 0}{N} \beta_i p_i (x)$$

In GSI $p_0 = 1$ a constant offset bias correction term.  The following predictors are:  the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight , and the emissivity sensitivity. Scan angle dependent bias is modeled as a 4th-order polynomial [Zhu 2014].

The $\beta_i$ coefficients are usually trained to achieve the best fit between the simulation and the observations. In the GSI system these coefficients are trained using a variational estimation method (Zhu 2014) which solves for the $\beta_i$ that provides the best fit between the simulation and the observations. The coefficients initialized from the 18 UTC Nov, 11 2018 GFS coefficients were trained during one week of continuous assimilation and using the same setup as in the experiments described in the next section. Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments to the predictor coefficients during each cycle.

During the quality control step, GSI rejects all observations if the innovation is greater than a predefined threshold for each channel and sensor. For microwave sensors, the quality control routines also use the scattering and Liquid Water Path (LWP) indexes calculated from brightness temperature channel differences (Weston et al., 2019; Zhu et al., 2016). The quality control routine for infrared sensors looks for observations over water with a large zenithal angle (over 60°) to reject channels near the visible range that can be contaminated with reflection. It also checks for cloud-affected observations using the transmittance profile calculated within the CRTM algorithms and performs an emissivity check for observations over land. In the experiment, only clear-sky observations are used as defined by the internal subroutines of GSI that use the LWP for microwave and the transmittance profile for infrared observations that could be potentially affected by clouds.

purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
  .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
  dcast(bufr_code ~ exp, value.var = "porcentaje") %>% 
  .[order(.[,'bufr_code']), ] %>% 
  kbl(booktabs = T, col.names = c("Obs type", "CONV", "AUT", "SATWND"),
      caption = "Observation types assimilated in each experiment.") %>% 
  kable_classic(full_width = F)

# purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
#   # .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
#   dcast(bufr_code ~ exp, value.var = "N") %>% 
#   .[order(.[,'bufr_code']), ] %>% 
#   kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND")) %>% 
#   kable_classic_2(full_width = F)
```

```{r}
fread(here("analysis", "data", "derived_data", "tabla_radianzas.csv")) %>% 
  .[, ":="(sensor = toupper(sensor),
           plataforma = toupper(plataforma),
           prop = scales::percent(prop/100, accuracy = 0.01, trim = TRUE))] %>% 
  .[order(.[,'sensor']), ] %>% 
  .[] %>% 
  kbl(booktabs = T, col.names = c("Sensor", "Plataform", "Assimilated\nchannels", "Percentage over total"),
      caption = "List of the available sensors in several platforms, the number of accepted channels and the percentage of assimilated observations calculated over all radiance observations.") %>% 
  collapse_rows(1,valign = "top") %>% 
  # column_spec(1, width = "3cm", latex_valign = "p") %>% 
  kable_classic_2(full_width = F)
```

```{r obs}
tribble(
  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
  "PREPBUFR", "x", "x", "x", "x",
  "AUTSFC", " ", "x", "x", "x",
  "SATWND", " ", " ", "x", "x",
  "RAD", " ", " ", " ", "x",
) %>% 
  kbl(booktabs = T, col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
      caption = "Observation types assimilated in each experiment.") %>% 
  kable_classic_2(full_width = F)
```

```{r dist_obs, fig.cap="a) Horizontal spatial distribution of the mean available observations per analysis cycle for CONV, AUT and SATWND. b) Number of radiance observations assimilated at each analysis cycle and pressure level."}
files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E2/ANA/*/diagfiles/asim*ensmean")[1:67]

@@

files <- files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67]

@@

files <- files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E6/ANA/*/diagfiles/asim*ensmean")[1:67]

@@

files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E8/ANA/*/diagfiles/asim*ensmean")

rad <- read_diag_rad(files, "E8") %>% 
  satinfo[., on = c("sensor", "channel")] %>% 
  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]

rbind(conv, aut, satwnd) %>% 
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, .(count_obs = .N), by = .(exp, lon.box, lat.box, date)] %>% 
  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lon.box, lat.box)] %>% 
  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
  geom_raster(aes(fill = obs_cycle), alpha = 0.8) +
  scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
                       guide = guide_colorbar(barwidth = 10,
                                              barheight = 0.5)) +
  geom_mapa() +
  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                               E6 = "SATWND", E8 = "RAD"))) +
  labs(fill = "Mean number of\nobs per cycle") +
  theme_minimal( base_size = 6) +
  theme(legend.position = "bottom",
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +

rad %>% 
  .[qc == 0 & iuse == 1] %>% 
  .[, lev := factor(cut_round(peakwt, breaks = seq(0, 1050, 50)))] %>% 
  .[, .(omb = mean(tbc),
        n = .N), by = .(date, lev, exp)] %>% 
  .[lev != 1025] %>% 
  ggplot(aes(date, fct_rev(lev))) +
  geom_tile(aes(fill = n)) +
  scale_fill_viridis_c(trans = scales::log10_trans(), breaks = c(1, 10, 100, 1000, 10000),
                       guide = guide_colourbar(barheight = 8,
                                               barwidth = 0.5)) +

  scale_date() +
   facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                               E6 = "SATWND", E8 = "RAD"))) +
  labs(x = NULL, fill = NULL, y = "pressure (hPa)") +
  theme_minimal(base_size = 6) +
  theme(aspect.ratio = 1/1.5) +

  plot_layout(ncol = 2, widths = c(1.6, 1)) + plot_annotation(tag_levels = 'a', tag_suffix = ")") &
  theme(plot.tag = element_text(size = 6))
```

```{r count_obs, fig.cap="a) Number of available observations per cycle during the experiment. a) Vertical spatial distribution of the mean available observations per analysis cycle."}
rad2 <- rad[, .(press, iuse, error, exp, date)] %>% 
  setnames(c("press", "iuse", "error"), c("pressure", "usage.flag", "rerr"))

rbind(conv, aut, satwnd) %>%
  .[, .(pressure, usage.flag, rerr, exp, date)] %>% 
  rbind(., rad2) %>% 
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, .(count_obs = .N), by = .(exp, date)] %>% 
  ggplot(aes(date, count_obs)) +
  geom_line(aes(color = exp), size = 0.2) +
  geom_point(aes(color = exp, shape = exp), size = 1) +
  scale_color_manual(values = colores_exp, 
                     labels = c(E2 = "CONV", E5 = "AUT", 
                                E6 = "SATWND", E8 = "RAD")) +
  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "AUT", 
                                                        E6 = "SATWND", E8 = "RAD")) +
  scale_y_log10() +
  scale_date() +
  labs(y = "Number of obs per cycle", x = NULL,
       color = NULL, shape = NULL) +
  theme_minimal(base_size = 6) +
  theme(legend.position = "bottom", 
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
